---
title: "CDS Spatial Analytics Assignment 2"
author: "Johan Horsmans"
output: github_document
---

# EXERCISE 1.1

```{r include = FALSE}
library(sf)
library(raster)
library(tmap)
library(spData)
library(spDataLarge)
```

#### The original map with my comments:
```{r}

tm_shape(nz_elev)  + # Plot elevation raster.
  tm_raster(title = "elev", # Title for elevation legend
            style = "cont", # Plot the scale continuously 
            palette = "BuGn") + # Color pallette
  tm_shape(nz) + # Plot polygon with the 16 regions of New Zealand.
  tm_borders(col = "red", # Color for the polygon lines.
             lwd = 3) + # With for the polygon lines.
  tm_scale_bar(breaks = c(0, 100, 200), # The breaks of the scale bar.
               text.size = 1) + # Text size for scale breaks.
  tm_compass(position = c("LEFT", "center"), # Create a compass + position.
             type = "rose", # Compass type.
             size = 2) + # Compass size.
  tm_credits(text = "A. Sobotkova, 2020") + # Credits name
  tm_layout(main.title = "My map", # Map title
            bg.color = "orange", # Background color
            inner.margins = c(0, 0, 0, 0)) # Margins (crop)
```


#### 1. Change the map title from "My map" to "New Zealand".

NOTE: Henceforth I will only comment the lines of code that are modified. For explanations of the other lines see the first chunk. 
```{r}
tm_shape(nz_elev)  +
  tm_raster(title = "elev", 
            style = "cont",
            palette = "BuGn") +
  tm_shape(nz) +
  tm_borders(col = "red", 
             lwd = 3) +
  tm_scale_bar(breaks = c(0, 100, 200),
               text.size = 1) +
  tm_compass(position = c("LEFT", "center"),
             type = "rose", 
             size = 2) +
  tm_credits(text = "A. Sobotkova, 2020") +
  tm_layout(main.title = "New Zealand", # Modified in order to change map title.
            bg.color = "orange",
            inner.margins = c(0, 0, 0, 0))
```

#### 2. Update the map credits with your own name and today's date.
```{r}
tm_shape(nz_elev)  +
  tm_raster(title = "elev", 
            style = "cont",
            palette = "BuGn") +
  tm_shape(nz) +
  tm_borders(col = "red", 
             lwd = 3) +
  tm_scale_bar(breaks = c(0, 100, 200),
               text.size = 1) +
  tm_compass(position = c("LEFT", "center"),
             type = "rose", 
             size = 2) +
  tm_credits(text = "Johan Horsmans, 11/02/21") + # Modified to change map credits.
  tm_layout(main.title = "New Zealand",
            bg.color = "orange",
            inner.margins = c(0, 0, 0, 0))
```

#### 3. Change the color palette to "-RdYlGn". 
```{r}
tm_shape(nz_elev)  +
  tm_raster(title = "elev", 
            style = "cont",
            palette = "-RdYlGn") + # Modified to change color palette.
  tm_shape(nz) +
  tm_borders(col = "red", 
             lwd = 3) +
  tm_scale_bar(breaks = c(0, 100, 200),
               text.size = 1) +
  tm_compass(position = c("LEFT", "center"),
             type = "rose", 
             size = 2) +
  tm_credits(text = "Johan Horsmans, 11/02/21") +
  tm_layout(main.title = "New Zealand",
            bg.color = "orange",
            inner.margins = c(0, 0, 0, 0))
```

#### 4. Put the north arrow in the top right corner of the map.
```{r}
tm_shape(nz_elev)  +
  tm_raster(title = "elev", 
            style = "cont",
            palette = "-RdYlGn") +
  tm_shape(nz) +
  tm_borders(col = "red", 
             lwd = 3) +
  tm_scale_bar(breaks = c(0, 100, 200),
               text.size = 1) +
  tm_compass(position = c("RIGHT", "top"), # Modifed to change the location of the compass.
             type = "rose", 
             size = 2) +
  tm_credits(text = "Johan Horsmans, 11/02/21") +
  tm_layout(main.title = "New Zealand",
            bg.color = "orange",
            inner.margins = c(0, 0, 0, 0))
```


#### 5. Improve the legend title by adding the used units (m asl).
```{r}
tm_shape(nz_elev)  +
  tm_raster(title = "Meters above sea level", # Modified to change the legend title. 
            style = "cont",
            palette = "-RdYlGn") +
  tm_shape(nz) +
  tm_borders(col = "red", 
             lwd = 3) +
  tm_scale_bar(breaks = c(0, 100, 200),
               text.size = 1) +
  tm_compass(position = c("RIGHT", "top"), 
             type = "rose", 
             size = 2) +
  tm_credits(text = "Johan Horsmans, 11/02/21") +
  tm_layout(main.title = "New Zealand",
            bg.color = "orange",
            inner.margins = c(0, 0, 0, 0))
```

#### 6. Increase the number of breaks in the scale bar.
```{r}
tm_shape(nz_elev)  +
  tm_raster(title = "Meters above sea level", 
            style = "cont",
            palette = "-RdYlGn") +
  tm_shape(nz) +
  tm_borders(col = "red", 
             lwd = 3) +
  tm_scale_bar(breaks = c(0, 50, 100, 150, 200), # Modified to increase the scale breaks.
               text.size = 1) +
  tm_compass(position = c("RIGHT", "top"), 
             type = "rose", 
             size = 2) +
  tm_credits(text = "Johan Horsmans, 11/02/21") +
  tm_layout(main.title = "New Zealand",
            bg.color = "orange",
            inner.margins = c(0, 0, 0, 0))
```

#### 7. Change the borders' color of the New Zealand's regions to black.
####    Decrease the line width.

```{r}
tm_shape(nz_elev)  +
  tm_raster(title = "Meters above sea level", 
            style = "cont",
            palette = "-RdYlGn") +
  tm_shape(nz) +
  tm_borders(col = "black", # Modified to make polygon lines black.
             lwd = 1) + # Modifeid to make the polygon lines thinner.
  tm_scale_bar(breaks = c(0, 50, 100, 150, 200), 
               text.size = 1) +
  tm_compass(position = c("RIGHT", "top"), 
             type = "rose", 
             size = 2) +
  tm_credits(text = "Johan Horsmans, 11/02/21") +
  tm_layout(main.title = "New Zealand",
            bg.color = "orange",
            inner.margins = c(0, 0, 0, 0))
```


#### 8. Change the background color to any color of your choice.
```{r}
tm_shape(nz_elev)  +
  tm_raster(title = "Meters above sea level", 
            style = "cont",
            palette = "-RdYlGn") +
  tm_shape(nz) +
  tm_borders(col = "black",  
             lwd = 1) + 
  tm_scale_bar(breaks = c(0, 50, 100, 150, 200), 
               text.size = 1) +
  tm_compass(position = c("RIGHT", "top"), 
             type = "rose", 
             size = 2) +
  tm_credits(text = "Johan Horsmans, 11/02/21") +
  tm_layout(main.title = "New Zealand",
            bg.color = "light blue", # Modified to make the background colour light blue.
            inner.margins = c(0, 0, 0, 0))
```
# EXERCISE 1.2

#### 9. Read two new datasets, `srtm` and `zion`, using the code below.

```{r}
srtm = raster(system.file("raster/srtm.tif", package = "spDataLarge")) # Read srtm (raster).
zion = read_sf(system.file("vector/zion.gpkg", package = "spDataLarge")) # Read zion (polygon).
```

####    To create a new map representing these datasets.

SRTM:
```{r}
tm_shape(srtm)  +
  tm_raster(title = "Meters above sea level", # Since it is a raster-object I use tm_raster to plot.
            style = "cont", # Continuous scale for the raster values.
            palette = "-RdYlGn") + # Color palette. 
  tm_credits(text = "Johan Horsmans, 11/02/21") + # Credits
  tm_layout(main.title = "SRTM", # Map Title
            inner.margins = c(0, 0, 0, 0)) # Margins
```

Zion:
```{r}
tm_shape(zion) + 
  tm_borders(col = "black")  +  # Make polygon lines black.
  tm_polygons() + # Since it is a vector-polygon-file I used tm_polygons to plot. 
  tm_credits(text = "Johan Horsmans, 11/02/21", position = c("RIGHT", "top")) + # Credits (placed in top right).
  tm_layout(main.title = "Zion", # Map title.
            bg.color = "light blue") # Background color. 

```

# EXERCISE 2.1:

```{r include = FALSE}
library(sf)
library(raster)
library(tmap)
library(spData)
library(spDataLarge)
```

#### 1. Display the `zion` object and view its structure.
```{r}
head(zion)
```
#### What can you say about the content of this file?
It is a "simple feature collection with 1 feature and 11 fields"

#### What type of data does it store? + how many attributes does it contain?

```{r}
plot(zion, max.plot = 11)
```

It stores the following variables/attributes: "Unit_code", "GIS_notes", "Unit_Name", "Date_edit", "State", "Region", "GNIS_ID", "Unit_type", "Created_by", "Metadata", "Parkname", "Geom"

#### What is the coordinate system used?
```{r}
st_crs(zion)
```
See information above.

#### What is its geometry?
The geometry type of Zion is a Polygon (see the head() output).


#### 2. Display the `srtm` object and view its structure.
#### What can you say about the content of this file? + What type of data does it store?

```{r}
srtm
```

```{r}
hist(getValues(srtm))
```

It is a raster and it contains values for each pixel indicating a certain parameter. The values of these pixels are distributed as in the plot shown above.

# What is the coordinate system used? 
```{r}
st_crs(srtm)
```
See above information.

#### How many attributes does it contain?
Only 1 since "attributes" are not specified in the first output.

#### How many dimensions does it have? 
457, 465, 212505  (nrow, ncol, ncell) (see first output).

#### What is the data resolution?
0.0008333333, 0.0008333333  (x, y) (see first output).


# Exercise 2.2

#### 1. Reproject the `srtm` dataset into the coordinate reference system used in the `zion` object. 
#### Create a new object `srtm2`
#### Vizualize the results using the `plot()` function.


```{r}
the_crs <- crs(zion, asText = TRUE) # Saving the coordinate reference system for the Zion object as "the_crs".

srtm2 <- projectRaster(srtm, crs = the_crs, method = "ngb") # Project the "srtm"-dataset into the zion-crs and save it as "srtm2".
```

Plotting srtm with the old and new CRS
```{r}
plot(srtm, main = "Original CRS")
#Versus
plot(srtm2, main = "New CRS")
```

As we see in the above output. The new CRS is slightly tilted towards the right and the value of the axes change. To illustrate that the two objects are correctly aligned, I will plot the zion-object on top of the alligned srtm-object:

```{r}
tm_shape(srtm2) + 
    tm_raster(title = "Meters above sea level", 
            style = "cont",
            palette = "-RdYlGn") + 
    tm_shape(zion) + 
    tm_polygons(alpha = 0.5)
```


As we can see here, they are sucessfully aligned.

#### 2. Reproject the `zion` dataset into the coordinate reference system used in the `srtm` object.
#### Create a new object `zion2`
#### Vizualize the results using the `plot()` function.
```{r}
the_crs <- crs(srtm, asText = TRUE) # Saving the coordinate reference system for the srtm object as "the_crs".

zion2 <- st_transform(zion, crs = the_crs) # Project the "zion"-dataset into the srtm-crs and save it as "zion2".
```

Plotting zion with the old and new CRS. For illustrative purposes I only plot a single attribute, namely "STATE".
```{r}
plot(zion["STATE"], main = "Old CRS") 
#Versus
plot(zion2["STATE"], main = "New CRS")
```


As we see in the above output. The new CRS is slightly tilted towards the left. To illustrate that the two objects are correctly aligned, I will plot the alligned zion-object on top of the srtm-object:

```{r}
#On top of each other
tm_shape(srtm) + 
    tm_raster(title = "Meters above sea level", 
            style = "cont",
            palette = "-RdYlGn") + 
    tm_shape(zion2) + 
    tm_polygons(alpha = 0.5)

```


As we can see here, they are sucessfully aligned.

